---
title: "Symbiont_Growth_Rates"
author: "Cheryl Logan"
date: "2/9/2021"
output: html_document
---

Load libraries
```{r}
library(tidyverse)
library(reshape2)
```

Define Symbiont Model Parameters from paper
```{r setup, include=FALSE}
# Eppley fxn parameters
a = 1.0768/12   # yr-1 - Linear Growth Rate (divide by 12 to get per month)
b = 0.0633       # C-1 - Exponential growth function
# evolutionary model parameters
selVx = 3.4627  # selectional variance
vg = .0025      # genetic variance (Zero when E=0!)
EnvVx = 0.0114  # environmental variance
# initialization for symbiont's optiomal temperature
g = 25         # initial temperature to which heat sensitive symbiont is optimally adapted
advantage = 1   # heat-tolerant symbiont advantage over heat sensitive symbiont (in degC)
g2 = g + advantage # temperature to which heat tolerant symbiont is optimally adapted
 
```

Get Eppley function values for each temperature between 0-35C
```{r}
# Create temp vector
t <- 1:35 # temp

# Fine Eppley function
rm <- matrix(nrow=35, ncol=3)
for(i in 1:35){
  rm[i,1] = a*exp(b*t(i))
  rm[i,2] = (a*exp(b*t(i)))/2
  rm[i,3] = (log(rm[i,2]/a))/b
}

df <- data.frame("Temp" = 1:35, "r_hat" = rm)
mdf <- reshape2::melt(df, id.var = "Temp")
```

Plot Eppley Function 
```{r}
ggplot(mdf, aes(x=Temp, y=value, colour=variable)) +
  geom_point() +
  geom_line() +
  ylim(-10,20) +
  xlab("Temperature (degC)") +
  ylab("Maximum Symbiont Growth Rate") 
```

Symbiont Growth curve

The "E221" curve:
ri(i,:) = (1- (vgi(i,:) + con.EnvVx + (min(2, gi(i,:) - temp(i))).^2) ./ (2*SelVx)) .* exp(con.b*min(0, temp(i) - gi(i,:) + 2)) * rm;

Written more simply:
mainpart = 1 - (vg +EnvVx + min(M1,g-T)^2 / (selVx)
extraExp = exp(X * b*min(0, T - g + M2))
r = mainpart * extraExp * rm


The values in mmx are
1) M1 The value in the first min function.
2) M2 The value in the "extra exponential", originally an argument to "min",
now a value added to gi.
3) X a multiplier on the "b" growth constant meant to steepen the curve
left of optimum.
4) B to select optional equation forms, currently the +1 advantage for shuffling.
5) 0/1 for default behavior or shifting max growth so that the tolerant symbiont 
gains no growth rate advantage.

Symbiont 1 = heat sensitive symbiont
Symbiont 2 = heat tolerant symbiont


```{r setup, include=TRUE}
# E221 Growth Curve
# note: for E221,SelVx cannot be below 2 without creating an error on the left side of the curve
# note: for E111,SelVx cannot be below .6 without creating an error on the left side of the curve
# note: for E001,SelVx can be anything
selVx = 2 # for small selVx values, the curve shape begins to increase on the cold side

M1 = 2  #abs(selVx/2) # minimum function in main part of eqn
M2 = 2  #abs(selVx/2) # extra exponential
X = 1

ri <- matrix(nrow=35, ncol=4)
for(T in 1:35){
  ri[T,1] = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx) 
  ri[T,2] = exp(X * b*min(0, T - g + M2))
  ri[T,3] = a*exp(b*T)
  ri[T,4] = ri[T,1] * ri[T,2] * ri[T,3]
}
df <- data.frame("Temp" = 1:35, "r_hat" = ri) # convert matrix to data frame

mdf <- reshape2::melt(df, id.var = "Temp")

#Plot Symbiont Growth Rate function
ggplot(df, aes(Temp,ri[,4])) +
  geom_point() +
  geom_line() +
  ylim(-.5, 0.6) +
  xlim(0, 39) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") 
```


#Marissa's original curve
```{r}
#Marissa's original curve
ri <- matrix(nrow=35, ncol=3)
for(T in 1:35){
  ri[T,1] = 1 - (vg + EnvVx + (g-T)^2) / (2*selVx) 
  ri[T,2] = a*exp(b*T)
  ri[T,3] = ri[T,1] * ri[T,2]
 }
 df <- data.frame("Temp" = 1:35, "r_hat" = ri) # convert matrix to data frame


 
#Marissa's original curve
ggplot(df, aes(x=Temp, y=ri[,3])) +
  geom_point() + 
  geom_line() +
  ylim(-2, 1) +
  xlim(15, 30) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  ggtitle( paste( "selVx =", selVx))

#what is T when ri is = (1 - (vg + EnvVx) / (2*selVx)) *a*exp(b*g)/2 ??
#when ri hits a*exp(b*g)/2 on cold side inflection begins
(a*exp(b*g)/2) /= 1 - (vg + EnvVx + (g-T)^2) / (2*selVx) 

```

#Try plotting multiple growth curves with same SelVx
```{r setup, include=TRUE}
X = 1
Temp = seq(1, 35, by=.1)  
selVx = 1
# for small selVx values, the curve shape begins to increase on the cold side
#SelV_vec <- c(.5, 1, 2, 4, 6, 8)
ri <- matrix(nrow=length(Temp), ncol=4)
  
cnt = 1
for(T in Temp) {
    
    M1 = 2; M2 = M1 #E221
    main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx) 
    extraexp = exp(X * b*min(0, T - g + M2))
    rm = a*exp(b*T)
    ri[cnt,1] = main * extraexp * rm
  
    #M1 = sqrt(abs(selVx-vg-EnvVx)); M2 = M1 #CL_new #.25*rmax
    #main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx) 
    #extraexp = exp(X * b*min(0, T - g + M2))
    #ri[cnt,2] = main * extraexp * rm
    
    M1 = sqrt(abs(1.5*selVx-vg-EnvVx)); M2 = M1 #CL_new #.25*rmax
    main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx) 
    extraexp = exp(X * b*min(0, T - g + M2))
    ri[cnt,3] = main * extraexp * rm
    
    M1 = sqrt(abs(1.8*selVx-vg-EnvVx)); M2 = M1 #CL_new #.25*rmax
    main = 1 - (vg + EnvVx + min(M1,(g+1)-T)^2) / (2*selVx) 
    extraexp = exp(X * b*min(0, T - (g+1) + M2))
    ri[cnt,4] = main * extraexp * rm
    
    cnt = cnt+1
  }

df <- data.frame("Temp" = Temp, "ri" = ri) # convert matrix to data frame
df <- rename(df, E221= ri.1, CL_new= ri.2, CL_new1.5= ri.3, CL_new_adv= ri.4)
#df <- rename(df, E221= ri.1, E111= ri.2, E001= ri.3, CL_new= ri.4, CL_v2= ri.5 )
#Curves <- c(E221, E111, E001, CL_new, CL_v2)SelVx=
mdf <- reshape2::melt(df, id.var = "Temp")

#Plot Symbiont Growth Rate function
ggplot(mdf, aes(x=Temp, y=value, colour=variable)) +
  geom_point(size = 0.1) + 
  geom_line() +
  ylim(-.2, .6) +
  xlim(15, 30) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  ggtitle( paste( "selVx =", selVx))
```

Test what M looks like for different values of SelV
```{r}
SelV_vec <- seq(.1, 8, by=.1)  
M <- matrix(nrow=length(SelV_vec), ncol=2)
cnt = 1 
for(selVx in SelV_vec) {
  M[cnt,1] = sqrt(abs(2*selVx-vg-EnvVx));
  M[cnt,2] = sqrt(abs(selVx-vg-EnvVx));
  cnt = cnt+1
}
df <- data.frame(M) # convert matrix to data frame


#Plot Symbiont Growth Rate function
ggplot(df, aes(x=SelV_vec)) +
  geom_point(aes(y=df[,1])) + 
  geom_point(aes(y=df[,2]), color='darkblue') + 
  ylim(0, 3) +
  xlim(0, 8) +
  xlab("SelVx") +
  ylab("M") +
  theme(legend.position="right") +
 ggtitle("M=CL_new")

```
