---
title: "Symbiont_Growth_Rates"
author: "Cheryl Logan"
date: "2/9/2021"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Load libraries
```{r load libraries, include=FALSE}
library(tidyverse)
library(reshape2)
```

Define Symbiont Model Parameters from Logan et al. 2021 paper
```{r params, include=FALSE}
# Eppley fxn parameters
a = 1.0768/12   # yr-1 - Linear Growth Rate (divide by 12 to get per month)
b = 0.0633       # C-1 - Exponential growth function
# evolutionary model parameters
selVx = 3.4627  # selectional variance
vg = .0025      # genetic variance (Zero when E=0!)
EnvVx = 0.0114  # environmental variance
# initialization for symbiont's optiomal temperature
g = 25         # initial temperature to which heat sensitive symbiont is optimally adapted
advantage = 1   # heat-tolerant symbiont advantage over heat sensitive symbiont (in degC)
g2 = g + advantage # temperature to which heat tolerant symbiont is optimally adapted
```


Symbiont Growth curve

The "E221" curve:
ri(i,:) = (1- (vgi(i,:) + con.EnvVx + (min(2, gi(i,:) - temp(i))).^2) ./ (2*SelVx)) .* exp(con.b*min(0, temp(i) - gi(i,:) + 2)) * rm;

Written more simply:
mainpart = 1 - (vg +EnvVx + min(M1,g-T)^2 / (selVx)
extraExp = exp(X * b*min(0, T - g + M2))
r = mainpart * extraExp * rm


The values in mmx are
1) M1 The value in the first min function.
2) M2 The value in the "extra exponential", originally an argument to "min",
now a value added to gi.
3) X a multiplier on the "b" growth constant meant to steepen the curve
left of optimum.
4) B to select optional equation forms, currently the +1 advantage for shuffling.
5) 0/1 for default behavior or shifting max growth so that the tolerant symbiont 
gains no growth rate advantage.

Symbiont 1 = heat sensitive symbiont
Symbiont 2 = heat tolerant symbiont



# Plot multiple growth curves with same selectional variance value (SelVx)
```{r Sym1 sym2, echo=FALSE}

X = -2
ri.adjust.1 = 2.6
ri.adjust.2 = 2.6

selVx=2

Temp = seq(1, 35, by=.1)  
# for small selVx values, the curve shape begins to increase on the cold side
#SelV_vec <- c(.5, 1, 2, 4, 6, 8)
ri <- matrix(nrow=length(Temp), ncol=2)
  
cnt = 1
for(T in Temp) {
    rm = a*exp(b*T)
    
    # M1 = 2; M2 = M1 #E221
    # main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx)
    # extraexp = exp(1 * b*min(0, T - g + M2))
    # ri[cnt,1] = main * extraexp * rm
    # 
    # M1 = 2; M2 = M1 #E221
    # main = 1 - (vg + EnvVx + min(M1,g2-T)^2) / (2*selVx)
    # extraexp = exp(1 * b*min(0, T - g2 + M2))
    # ri[cnt,2] = main * extraexp * rm
    # # 
    # M1 = sqrt(abs(ri.adjust.1*selVx-vg-EnvVx)); M2 = M1 #CL_new #.25*rmax
    # main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx)
    # extraexp = exp(X * b*min(0, T - g + M2))
    # ri[cnt,3] = main * extraexp * rm
    # 
    # M1 = sqrt(abs(ri.adjust.2*selVx-vg-EnvVx)); M2 = M1 #CL_new #.25*rmax
    # main = 1 - (vg + EnvVx + min(M1,g2-T)^2) / (2*selVx)
    # extraexp = exp(X * 1*b*min(0, T - g2 + M2))
    # ri[cnt,4] = main * extraexp * rm
    
    M1 = sqrt(abs(ri.adjust.1*selVx)); M2 = M1 #simple
    main = 1 - (vg + EnvVx + min(M1,g-T)^2) / (2*selVx)
    extraexp = exp(X * b*min(0, T - g + M2))
    ri[cnt,1] = main * extraexp * rm

    M1 = sqrt(abs(ri.adjust.2*selVx)); M2 = M1 #simple
    main = 1 - (vg + EnvVx + min(M1,g2-T)^2) / (2*selVx)
    extraexp = exp(X * 1*b*min(0, T - g2 + M2))
    ri[cnt,2] = main * extraexp * rm
    
    cnt = cnt+1
  }

df <- data.frame("Temp" = Temp, "ri" = ri) # convert matrix to data frame
df <- rename(df, symbiont_1_new= ri.1, symbiont_2_new= ri.2)
#df <- rename(df, E221= ri.1, E111= ri.2, E001= ri.3, CL_new= ri.4, CL_v2= ri.5 )
#Curves <- c(E221, E111, E001, CL_new, CL_v2)SelVx=
mdf <- reshape2::melt(df, id.var = "Temp")

#Plot Symbiont Growth Rate function
cols <- c("symbiont_1_new" = "black", "symbiont_2_new" = "blue")

ggplot(mdf, aes(x=Temp, y=value, colour=variable)) +
  geom_point(size = 0.1) + 
  geom_line(size = 0.9) +
  ylim(-.3, .5) +
  xlim(15, 30) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  theme_classic() +
  scale_colour_manual(values = cols) +
  ggtitle( paste( "selVx=", selVx, ", sym1adj=",ri.adjust.1 , ", sym2adj=",ri.adjust.2, ", X=", X ) )

ggsave("growth_curve_new.png", height = 3, width = 5)
```




# Plot growth curve across range of  SelVx
```{r SelV Range, echo=FALSE}
X = -2
Temp = seq(1, 35, by=.1)  
SelV_vec <- c(.25, .5, 1, 2, 4, 6, 8)
ri <- matrix(nrow=length(Temp), ncol=length(SelV_vec))
ri.adjust = 2.6

Scnt = 1
for(selVx in SelV_vec) {
  cnt = 1
  for (T in Temp) {
    M1 = sqrt((ri.adjust * selVx - vg - EnvVx)); M2 = M1
    rm = a*exp(b*T)
    main = 1 - (vg + EnvVx + min(M1, g - T) ^ 2) / (2 * selVx)
    extraexp =  exp(X * b * min(0, T - g + M2))
    ri[cnt, Scnt] = main * extraexp * rm
    cnt = cnt + 1
  }
  Scnt = Scnt + 1
}

df <- data.frame("Temp" = Temp, "ri" = ri) # convert matrix to data frame
df <- rename(df, SelV_.25= ri.1, SelV_.5= ri.2, SelV_1= ri.3, SelV_2= ri.4, SelV_4= ri.5, SelV_6= ri.6, SelV_8= ri.7)
mdf <- reshape2::melt(df, id.var = "Temp")

ggplot(mdf, aes(x=Temp, y=value, colour=variable)) +
  geom_point(size = 0.1) + 
  geom_line() +
  ylim(-.5, .5) +
  xlim(10, 35) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  ggtitle( paste( "ri.adjust =", ri.adjust, ", X =", X))
```



Baskett et al. 2009 original sym growth curve
```{r echo=FALSE}
#Marissa's original curve
selVx = 2
Temp = seq(1, 35, by=.1)  
ri <- matrix(nrow=length(Temp), ncol=3)
cnt=1
for(T in Temp){
  ri[cnt,1] = 1 - (vg + EnvVx + (g-T)^2) / (2*selVx) 
  ri[cnt,2] = a*exp(b*T)
  ri[cnt,3] = ri[cnt,1] * ri[cnt,2]
  cnt = cnt+1  
}
 df <- data.frame("r_hat" = ri) # convert matrix to data frame

# plot original curve
ggplot(df, aes(x=Temp, y=ri[,3])) +
  geom_point() + 
  geom_line() +
  #ylim(-.5, .5) +
  #xlim(15, 30) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  ggtitle( paste( " Symbiont growth curve (Baskett et al. 2009), selVx =", selVx))
```
Baskett et al. 2009 original sym growth curve
```{r echo=FALSE}
#Marissa's original curve
selVx = 2
Temp = seq(1, 35, by=.1)  
ri <- matrix(nrow=length(Temp), ncol=3)
cnt=1
for(T in Temp){
  ri[cnt,1] = 1 - (vg + EnvVx + (g-T)^2) / (2*selVx) 
  ri[cnt,2] = a*exp(b*T)
  ri[cnt,3] = ri[cnt,1] * ri[cnt,2]
  cnt = cnt+1  
}
 df <- data.frame("r_hat" = ri) # convert matrix to data frame

# plot original curve
ggplot(df, aes(x=Temp, y=ri[,3])) +
  geom_point() + 
  geom_line() +
  #ylim(-.5, .5) +
  #xlim(15, 30) +
  xlab("Temperature (degC)") +
  ylab("Symbiont Growth Rate") +
  theme(legend.position="right") + 
  ggtitle( paste( " Symbiont growth curve (Baskett et al. 2009), selVx =", selVx))
```